Algorithms for Longest Common Extensions

### 2011-08-19 presentation/presentation.tex
m

\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{Existing Algorithm: \proc{DirectComp}}
    \begin{tabular}{r l}
        Preprocessing & $O(1)$ \\
        Space & $O(1)$ \\
        Query & $O(|\LCE(i,j)|)=O(n)$ \\
        Average query & $O(1)$ \\
        Query I/Os & $O\Big(\frac{|\LCE(i,j)|}{B}\Big)$ \\
    \end{tabular}

@@

\subsection{The \proc{LcpRmq} Algorithm}
\begin{frame}{Existing Algorithm: \proc{LcpRmq}}
    \begin{multicols}{2}{
        \begin{align*}
            s&=abbababba\\
            s[2\twodots n]&=bbababba\\
            s[3\twodots n]&=bababba\\
            \LCE_s(2,3)&=1
        \end{align*}
        \newpage
        \begin{center}
            %\uncover<2->{
            \includegraphics[width=0.26\textwidth,page=1]{../doc/sa+lcp+min.pdf}
            %}
        \end{center}
    }
    \end{multicols}
    %\uncover<3->{
    \begin{center}
        $\LCE(i,j)=\LCP[\RMQ_{\LCP}(\SAinv[i] + 1, \SAinv[j])]$\\
        where $\SAinv[i] < \SAinv[j]$\\
        \vspace{1em}
        \begin{tabular}{r l}
            Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$ \\
            Query & $O(1)$ \\
            Average query & $O(1)$ \\
            Query I/Os & $O(1)$ \\
        \end{tabular}
    \end{center}
    %}
\end{frame}

@@

\subsection{I/O}
\begin{frame}{The \fprintk\ Algorithm: I/O}
    \begin{itemize}
        \item Original:
        \begin{itemize}
            \item Data structure: $H_\ell[i] = F_{t_\ell}[i]$
            \item Size: $|H_\ell| = n$
            \item I/Os: $O\big(k\cdot n^{1/k}\big)$
        \end{itemize}
        \item Cache optimized:
        %Instead of storing the fingerprint of $s[i\twodots i+t_\ell-1]$ at $H_\ell[i]$, we can store it at 
        \begin{itemize}
            \item Data structure: \\~~$H_\ell[((i-1)\mod t_\ell)\cdot\lceil n/t_\ell\rceil+\lfloor (i-1)/t_\ell\rfloor+1] = F_{t_\ell}[i]$
            \item Size: $|H_\ell| = n+t_\ell$
            \item I/Os: $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$
            \begin{itemize}
                \item Best when $k$ is small $\implies$ $n^{1/k}$ is large.
            \end{itemize}
        \end{itemize}
     \end{itemize}
     %$n+t_\ell-1/t_\ell+2$
\end{frame}

\begin{frame}{The \fprintk\ Algorithm}
    \begin{tabular}{r l}
        Preprocessing & $O(k\cdot n + \sortt)$ \\
        Space & $O(k\cdot n)$\\
        Query & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ \\
        Query I/Os & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}

@@

        \begin{tabular}{r l}
            Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$\\
            Query & $O(\sqrt n)$ \\
            Average query & $O(1)$ \\
            Query I/Os & $O\Big(\frac{\sqrt n}{B}\Big)$ \\
        \end{tabular}
        \newpage
        $k=\logceil$:

        \begin{tabular}{r l}
            Preprocessing & $O(n\log n)$ \\
            Space & $O(n\log n)$\\
            Query & $O(\log n)$ \\
            Average query & $O(1)$ \\
            Query I/Os & $O(\log n)$ \\
        \end{tabular}
    }
    \end{multicols}
\end{frame}

@@

\subsection{Practical I/O}
\begin{frame}{The \fprintk\ Algorithm: Practical Results, I/O}
    Query times of \fprint[2]
    \textcolor{red}{without cache optimization}
    and with cache optimization using
    \textcolor{green}{shift operations} vs.\
    \textcolor{blue}{multiplication and division}
    %by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp2-rand10.plt}
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp2-alla.plt}
        \end{center}
    }
    \end{multicols}
\end {frame}

\subsection{Practical I/O}
\begin{frame}{The \fprintk\ Algorithm: Practical Results, I/O}
    Query times of \fprint[3]
    \textcolor{red}{without cache optimization}
    and with cache optimization using
    \textcolor{green}{shift operations} vs.\
    \textcolor{blue}{multiplication and division}
    %by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp3-rand10.plt}
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp3-alla.plt}
        \end{center}
    }
    \end{multicols}
\end {frame}

>>>>

\subsection{The \proc{DirectComp} algorithm}
\begin{frame}{Existing Algorithm: \proc{DirectComp}}
    \begin{tabular}{r l}
        Preprocessing & $O(1)$ \\
        Space & $O(1)$ \\
        Query & $O(|\LCE(i,j)|)=O(n)$ \\
        Average query & $O(1)$ \\
        Query I/O & $O\Big(\frac{|\LCE(i,j)|}{B}\Big)$ \\
    \end{tabular}

@@

\subsection{The \proc{LcpRmq} Algorithm}
\begin{frame}{Existing Algorithm: \proc{LcpRmq}}
    \begin{multicols}{2}{
        \begin{align*}
            s&=abbababba\\
            s[2\twodots n]&=bbababba\\
            s[3\twodots n]&=bababba\\
            \LCE_s(2,3)&=1
        \end{align*}
        \newpage
        \begin{center}
            %\uncover<2->{
            \includegraphics[width=0.26\textwidth,page=1]{../doc/sa+lcp+min.pdf}
            %}
        \end{center}
    }
    \end{multicols}
    %\uncover<3->{
    \begin{center}
        $\LCE(i,j)=\LCP[\RMQ_{\LCP}(\SAinv[i] + 1, \SAinv[j])]$\\
        where $\SAinv[i] < \SAinv[j]$\\
        \vspace{1em}
        \begin{tabular}{r l}
            Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$ \\
            Query & $O(1)$ \\
            Average query & $O(1)$ \\
            Query I/O & $O(1)$ \\
        \end{tabular}
    \end{center}
    %}
\end{frame}

@@

\subsection{I/O}
\begin{frame}{The \fprintk\ Algorithm: I/O}
    \begin{itemize}
        \item Original:
        \begin{itemize}
            \item Data structure: $H_\ell[i] = F_{t_\ell}[i]$
            \item Size: $|H_\ell| = n$
            \item I/O: $O\big(k\cdot n^{1/k}\big)$
        \end{itemize}
        \item Cache optimized:
        %Instead of storing the fingerprint of $s[i\twodots i+t_\ell-1]$ at $H_\ell[i]$, we can store it at 
        \begin{itemize}
            \item Data structure: \\~~$H_\ell[((i-1)\mod t_\ell)\cdot\lceil n/t_\ell\rceil+\lfloor (i-1)/t_\ell\rfloor+1] = F_{t_\ell}[i]$
            \item Size: $|H_\ell| = n+t_\ell$
            \item I/O: $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$
            \begin{itemize}
                \item Best when $k$ is small $\implies$ $n^{1/k}$ is large.
            \end{itemize}
        \end{itemize}
     \end{itemize}
     %$n+t_\ell-1/t_\ell+2$
\end{frame}

\begin{frame}{The \fprintk\ Algorithm}
    \begin{tabular}{r l}
        Preprocessing & $O(k\cdot n + \sortt)$ \\
        Space & $O(k\cdot n)$\\
        Query & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ \\
        Query I/O & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}

@@

        \begin{tabular}{r l}
            Preprocessing & $O(\sortt)$ \\
            Space & $O(n)$\\
            Query & $O(\sqrt n)$ \\
            Average query & $O(1)$ \\
            Query I/O & $O\Big(\frac{\sqrt n}{B}\Big)$ \\
        \end{tabular}
        \newpage
        $k=\logceil$:

        \begin{tabular}{r l}
            Preprocessing & $O(n\log n)$ \\
            Space & $O(n\log n)$\\
            Query & $O(\log n)$ \\
            Average query & $O(1)$ \\
            Query I/O & $O(\log n)$ \\
        \end{tabular}
    }
    \end{multicols}
\end{frame}

@@

%\subsection{Practical I/O}
\begin{frame}{The \fprintk\ Algorithm: Practical Results, I/O}
    Query times of \fprint[2]
    \textcolor{red}{without cache optimization}
    and with cache optimization using
    \textcolor{green}{shift operations} vs.\
    \textcolor{blue}{multiplication and division}
    %by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp2-rand10.plt}
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp2-alla.plt}
        \end{center}
    }
    \end{multicols}
\end{frame}

\begin{frame}{The \fprintk\ Algorithm: Practical Results, I/O}
    Query times of \fprint[3]
    \textcolor{red}{without cache optimization}
    and with cache optimization using
    \textcolor{green}{shift operations} vs.\
    \textcolor{blue}{multiplication and division}
    %by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp3-rand10.plt}
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-slides-cache-fp3-alla.plt}
        \end{center}
    }
    \end{multicols}
\end{frame}

\section{LCE on trees}
\subsection{Compression}
\begin{frame}{LCE on Compressed Strings}
    Goal
    \begin{itemize}
        \item Allow LCE queries without decompressing the string
        \item Using Ziv-Lempel compression (LZ)
    \end{itemize}
    How
    \begin{itemize}
        \item LZ compression represents the string as a tree
        \item An LCE query on a LZ compressed string is a number of LCE queries on a tree
    \end{itemize}
\end{frame}

\begin{frame}{LCE on Trees}
    \begin{itemize}
        \item Trees:
        \begin{itemize}
            \item One character on each edge
            \item LCE is the length of the longest common prefix of two strings along two paths
        \end{itemize}
        \item $p_i=bbc$ and $p_j=bbba$ gives $\LCE(p_i,p_j)=2$
    \end{itemize}
    \begin{center}
        \includegraphics[width=0.4\textwidth,page=1]{tree-lce.pdf}
    \end{center}
\end{frame}
